# R markdown for prediction1
## setup
```{r setup}
knitr::opts_knit$set(root.dir = 'C:/Users/אחיעד/Desktop/data science/ass2/prediction3')
getwd()
```

## read train.csv into dataframe
```{r}
df <- read.csv("C:/Users/אחיעד/Desktop/data science/ass2/Titanic/train.csv",na.strings = "")
```

## using str to identify data types
```{r}
str(df)
```

## convert Survived column and Pclass column to factors
```{r}
df$Survived<- as.factor(df$Survived)
df$Pclass<- as.factor(df$Pclass)
```

## data summary
```{r}
summary(df)
```

## remove features from df
```{r}
df <- df[,-c(1,4,9)]
```

## devide into factors and numerics columns
```{r}
cols<- 1:dim(df)[2]
factors <- cols[sapply(df,is.factor)]
numerics <- cols[!sapply(df,is.factor)]
```

## We now tide the data two times: the first is for categorial data and the second for numeric data.
```{r}
#install.packages("tidyr")
library(tidyr)
df_tidy_factors<-gather(df[,factors],"feature","value", -1)
df_tidy_numerics<-gather(cbind(Survived=df[,1],df[,numerics]),"feature","value",-1)

```

## plot for factor columns
```{r}
#install.packages("ggplot2")
library(ggplot2)
qplot(x=value,data=df_tidy_factors,fill=Survived) + facet_grid(~feature,scales="free")
```

## plot for numerical columns
```{r}
qplot(x=value,data=df_tidy_numerics,fill=Survived) + facet_grid(~feature,scales="free")
```

## we trained the rpart algorithm, using cross validation control on the data provided.
```{r}
#install.packages("caret")
#instail.packages("lattice")
#install.packages("gbm")
library(caret)
library(lattice)
set.seed(13)
control <- trainControl(method="cv", number=10)
fit.rpart <- train(Survived~., data=df, method="rpart", metric="Kappa", trControl=control,na.action = na.pass)
fit.rpart
```

## show graphical view of our model that show the accuracy according to the complexity parameter.
```{r}
plot(fit.rpart)
```

## load the test file, with na as ""
```{r}
new_df <-read.csv('C:/Users/אחיעד/Desktop/data science/ass2/Titanic/test.csv',na.strings = "")
```

## Create a vector with the PassengerIds of records in the test file. 
```{r}
ids<- new_df$PassengerId
```

## remove features from new_df(the test file)
```{r}
new_df$Pclass<- as.factor(new_df$Pclass)
new_df<- new_df[,-c(1,3,8)]
```


## Another thing that is going to make troubles is that the test data contains some new levels in the *Cabin* feature that did not appear in the train data (you can check it by applying *Summary* on the test data). Add these new levels to the model.

### Getting the levels of the "Cabin" feature from a model *m*: m$xlevels[["Cabin"]]

### Getting the levels of the "Cabin" feature from a dataframe *d*: levels(d$Cabin)

### Union operation: union(x,y)

```{r}
fit.rpart$xlevels[["Cabin"]] <- union(fit.rpart$xlevels[["Cabin"]], levels(new_df$Cabin))
```


## we predict the rpart model on the new_df data set, in the prediction process we pass the na fields in the data set, because the prediction does not succedd with na fields.
### this prediction is in order to predict the test.csv file. 
```{r}
new_pred_rpart<- predict(fit.rpart,new_df,na.action = na.pass)
```

### show how much passengers survived and died according to our model based on rpart algorhitm.
```{r}
plot(new_pred_rpart)
```

### Write the *PassengerId* and *Survived* attributes to a csv file.
```{r}
res <- cbind(PassengerId=ids,Survived=as.character(new_pred_rpart))
write.csv(res,file="C:/Users/אחיעד/Desktop/data science/ass2/prediction3/output.csv",row.names = F)
```