---
title: "prediction3"
output: html_document
---

#R markdown for prediction1
##setup
```{r setup}
knitr::opts_knit$set(root.dir = 'C:/Users/אחיעד/Desktop/data science/ass2/prediction3')
getwd()
```

##read train.csv into dataframe
```{r}
df <- read.csv("C:/Users/אחיעד/Desktop/data science/ass2/Titanic/train.csv",na.strings = "")
```

##using str to identify data types
```{r}
str(df)
```

```{r}
df$Survived<- as.factor(df$Survived)
df$Pclass<- as.factor(df$Pclass)
```

##data summary
```{r}
summary(df)
```
##remove features from df

```{r}
df <- df[,-c(1,4,9)]
```

```{r}
cols<- 1:dim(df)[2]
factors <- cols[sapply(df,is.factor)]
numerics <- cols[!sapply(df,is.factor)]
```

We now tide the data two times: the first is for categorial data and the second for numeric data.

```{r}
#install.packages("tidyr")
library(tidyr)
df_tidy_factors<-gather(df[,factors],"feature","value", -1)
df_tidy_numerics<-gather(cbind(Survived=df[,1],df[,numerics]),"feature","value",-1)

```


Finally, we can plot. The first plot describes only categorical features (factors). 
Notice that the *scales* parameter was set to "free" to enable a suitable scaling for each facet (otherwise it is hard to view some of the facets, that need much smaller scales). We use the *facet_grid* that accepts a *scales* parameter.

```{r}
#install.packages("ggplot2")
library(ggplot2)
qplot(x=value,data=df_tidy_factors,fill=Survived) + facet_grid(~feature,scales="free")
```


One more plot for numeric features:
```{r}
qplot(x=value,data=df_tidy_numerics,fill=Survived) + facet_grid(~feature,scales="free")
```

13. (Non-mandatory) Now train the same model using the caret package. This time set up a 10-fold cross validation configuration instead of the train-test split that was used before. (No need to set any parameters for the model, go with its default tuning grid).
Set the *na.action* to "na.pass" in order for caret to pass NAs to C5.0 instead of failing.
Check the model's average test accuracy, and the parameter values that gave the best accuracy.

```{r}
#install.packages("caret")
#instail.packages("lattice")
#install.packages("gbm")
library(caret)
library(lattice)
set.seed(13)
control <- trainControl(method="cv", number=10)
fit.rpart <- train(Survived~., data=df, method="rpart", metric="Kappa", trControl=control,na.action = na.pass)
fit.rpart
```

```{r}
plot(fit.rpart)
```
14. Load the test file. Don't forget the empty strings issue when loading the file (na.strings = "").

```{r}
new_df <-read.csv('C:/Users/אחיעד/Desktop/data science/ass2/Titanic/test.csv',na.strings = "")
```


15. Create a vector with the PassengerIds of records in the test file. We will soon attach it to the prediction results.

```{r}
ids<- new_df$PassengerId
```

16. Repeat the same preprocessing steps that were performed on the train data (factorizing two features and ignoring three features). You must make sure that the train and test data have the same structure (except for the target feature)

```{r}
new_df$Pclass<- as.factor(new_df$Pclass)
new_df<- new_df[,-c(1,3,8)]
```


17. Another thing that is going to make troubles is that the test data contains some new levels in the *Cabin* feature that did not appear in the train data (you can check it by applying *Summary* on the test data). Add these new levels to the model.

- Getting the levels of the "Cabin" feature from a model *m*: m$xlevels[["Cabin"]]

- Getting the levels of the "Cabin" feature from a dataframe *d*: levels(d$Cabin)

- Union operation: union(x,y)


```{r}
fit.rpart$xlevels[["Cabin"]] <- union(fit.rpart$xlevels[["Cabin"]], levels(new_df$Cabin))
```


18. Load the test file and predict the target attribute for its records using your trained model. Don't forget to set *na.action* to "na.pass" if you used caret in the model creation, otherwise it will fail.

```{r}
new_pred_rpart<- predict(fit.rpart,new_df,na.action = na.pass)
```


```{r}
plot(new_pred_rpart)
```

19. Write the *PassengerId* and *Survived* attributes to a csv file and submit this file to kaggle's competition (You should first subscribe to the site). What is your score?

```{r}
res <- cbind(PassengerId=ids,Survived=as.character(new_pred_rpart))
write.csv(res,file="C:/Users/אחיעד/Desktop/data science/ass2/prediction3/output.csv",row.names = F)
```



